### 2025-07-02 machine-learning.qmd
20250702 - clean up

# Set up cross-validation
folds <- rsample::group_vfold_cv(
  data_train_qb,
  group = gsis_id,
  v = 10) # 10-fold cross-validation

@@

# Fit Model with Cross-Validation
cv_results <- tune::fit_resamples(
  lm_wf,
  resamples = folds,
  metrics = metric_set(rmse, mae, rsq),
  control = control_resamples(save_pred = TRUE)
)

@@

# Predict on Test Data
predict(final_model, data_test_qb)

df <- data_test_qb %>%
  mutate(pred = predict(final_model, new_data = data_test_qb)$.pred)

@@

# Set up cross-validation
folds <- rsample::group_vfold_cv(
  data_train_qb,
  group = gsis_id,
  v = 10) # 10-fold cross-validation

# Define Recipe (Formula)
rec <- recipes::recipe(
  fantasyPoints_lag ~ .,
  data = data_train_qb %>% select(-gsis_id, -fantasyPointsMC_lag))

@@

# Set up cross-validation
folds <- rsample::group_vfold_cv(
  data_train_qb,
  group = gsis_id,
  v = 10) # 10-fold cross-validation

# Define Recipe (Formula)
rec <- recipes::recipe(
  fantasyPoints_lag ~ .,
  data = data_train_qb %>% select(-gsis_id, -fantasyPointsMC_lag))

# Define Model
lasso_spec <- 
  parsnip::linear_reg(
    penalty = tune(),
    mixture = 1) %>%
  set_engine("glmnet")

@@

# Tune the Penalty Parameter
cv_results <- tune::tune_grid(
  lasso_wf,
  resamples = folds,
  grid = penalty_grid,
  metrics = metric_set(rmse, mae, rsq),
  control = control_grid(save_pred = TRUE)
)

@@

# Identify best penalty
select_best(cv_results, metric = "rmse")
select_best(cv_results, metric = "mae")
select_best(cv_results, metric = "rsq")

best_penalty <- select_best(cv_results, metric = "mae")

# Finalize Workflow with Best Penalty
final_wf <- finalize_workflow(
  lasso_wf,
  best_penalty)

@@

# Set up cross-validation
folds <- rsample::group_vfold_cv(
  data_train_qb,
  group = gsis_id,
  v = 10) # 10-fold cross-validation

# Define Recipe (Formula)
rec <- recipes::recipe(
  fantasyPoints_lag ~ .,
  data = data_train_qb %>% select(-gsis_id, -fantasyPointsMC_lag))

# Define Model
ridge_spec <- 
  linear_reg(
    penalty = tune(),
    mixture = 0) %>%
  set_engine("glmnet")

@@

# Tune the Penalty Parameter
cv_results <- tune::tune_grid(
  ridge_wf,
  resamples = folds,
  grid = penalty_grid,
  metrics = metric_set(rmse, mae, rsq),
  control = control_grid(save_pred = TRUE)
)

@@

# Identify best penalty
select_best(cv_results, metric = "rmse")
select_best(cv_results, metric = "mae")
select_best(cv_results, metric = "rsq")

best_penalty <- select_best(cv_results, metric = "mae")

# Finalize Workflow with Best Penalty
final_wf <- finalize_workflow(
  ridge_wf,
  best_penalty)

@@

# Set up cross-validation
folds <- rsample::group_vfold_cv(
  data_train_qb,
  group = gsis_id,
  v = 10) # 10-fold cross-validation

# Define Recipe (Formula)
rec <- recipes::recipe(
  fantasyPoints_lag ~ .,
  data = data_train_qb %>% select(-gsis_id, -fantasyPointsMC_lag))

# Define Model
enet_spec <- 
  linear_reg(
    penalty = tune(),
    mixture = tune()) %>%
  set_engine("glmnet")

@@

# Tune the Grid
cv_results <- tune::tune_grid(
  enet_wf,
  resamples = folds,
  grid = grid_enet,
  metrics = metric_set(rmse, mae, rsq),
  control = control_grid(save_pred = TRUE)
)

@@

# Identify best penalty
select_best(cv_results, metric = "rmse")
select_best(cv_results, metric = "mae")
select_best(cv_results, metric = "rsq")

best_penalty <- select_best(cv_results, metric = "mae")

# Finalize Workflow with Best Penalty
final_wf <- finalize_workflow(
  enet_wf,
  best_penalty)

@@

# Set up cross-validation
folds <- rsample::group_vfold_cv(
  data_train_qb,
  group = gsis_id,
  v = 10) # 10-fold cross-validation

# Define Recipe (Formula)
rec <- recipes::recipe(
  fantasyPoints_lag ~ .,
  data = data_train_qb %>% select(-gsis_id, -fantasyPointsMC_lag))

# Define Model
rf_spec <- 
  parsnip::rand_forest(
    mtry = tune::tune(),
    min_n = tune::tune(),
    trees = 500) %>%
  parsnip::set_mode("regression") %>%
  parsnip::set_engine("ranger", importance = "impurity")

@@

# Tune the Grid
cv_results <- tune::tune_grid(
  rf_wf,
  resamples = folds,
  grid = rf_grid,
  metrics = metric_set(rmse, mae, rsq),
  control = control_grid(save_pred = TRUE)
)

@@

# Identify best penalty
select_best(cv_results, metric = "rmse")
select_best(cv_results, metric = "mae")
select_best(cv_results, metric = "rsq")

best_penalty <- select_best(cv_results, metric = "mae")

# Finalize Workflow with Best Penalty
final_wf <- finalize_workflow(
  rf_wf,
  best_penalty)

@@

smerf$forest # the fitted random forest (obtained at the last iteration)
smerf$random_effects # the predicted random effects for each player
smerf$omega # the predicted stochastic processes
smerf$OOB # OOB error at each iteration
```

>>>>

# Set up Cross-Validation
folds <- rsample::group_vfold_cv(
  data_train_qb,
  group = gsis_id, # ensures all rows for a player are in the training set or all in the validation set for each fold
  v = 10) # 10-fold cross-validation

@@

# Fit Model with Cross-Validation
cv_results <- tune::fit_resamples(
  lm_wf,
  resamples = folds,
  metrics = yardstick::metric_set(rmse, mae, rsq),
  control = tune::control_resamples(save_pred = TRUE)
)

@@

# Predict on Test Data
df <- data_test_qb %>%
  mutate(pred = predict(final_model, new_data = data_test_qb)$.pred)

@@

# Set up Cross-Validation
folds <- rsample::group_vfold_cv(
  data_train_qb,
  group = gsis_id, # ensures all rows for a player are in the training set or all in the validation set for each fold
  v = 10) # 10-fold cross-validation

# Define Recipe (Formula)
rec <- recipes::recipe(
  fantasyPoints_lag ~ ., # use all predictors
  data = data_train_qb %>% select(-gsis_id, -fantasyPointsMC_lag))

@@

# Set up Cross-Validation
folds <- rsample::group_vfold_cv(
  data_train_qb,
  group = gsis_id, # ensures all rows for a player are in the training set or all in the validation set for each fold
  v = 10) # 10-fold cross-validation

# Define Recipe (Formula)
rec <- recipes::recipe(
  fantasyPoints_lag ~ ., # use all predictors
  data = data_train_qb %>% select(-gsis_id, -fantasyPointsMC_lag))

# Define Model
lasso_spec <- 
  parsnip::linear_reg(
    penalty = tune::tune(),
    mixture = 1) %>%
  parsnip::set_engine("glmnet")

@@

# Tune the Penalty Parameter
cv_results <- tune::tune_grid(
  lasso_wf,
  resamples = folds,
  grid = penalty_grid,
  metrics = yardstick::metric_set(rmse, mae, rsq),
  control = tune::control_grid(save_pred = TRUE)
)

@@

# Identify best penalty
tune::select_best(cv_results, metric = "rmse")
tune::select_best(cv_results, metric = "mae")
tune::select_best(cv_results, metric = "rsq")

best_penalty <- tune::select_best(cv_results, metric = "mae")

# Finalize Workflow with Best Penalty
final_wf <- tune::finalize_workflow(
  lasso_wf,
  best_penalty)

@@

# Set up Cross-Validation
folds <- rsample::group_vfold_cv(
  data_train_qb,
  group = gsis_id, # ensures all rows for a player are in the training set or all in the validation set for each fold
  v = 10) # 10-fold cross-validation

# Define Recipe (Formula)
rec <- recipes::recipe(
  fantasyPoints_lag ~ ., # use all predictors
  data = data_train_qb %>% select(-gsis_id, -fantasyPointsMC_lag))

# Define Model
ridge_spec <- 
  parnsnip::linear_reg(
    penalty = tune::tune(),
    mixture = 0) %>%
  parsnip::set_engine("glmnet")

@@

# Tune the Penalty Parameter
cv_results <- tune::tune_grid(
  ridge_wf,
  resamples = folds,
  grid = penalty_grid,
  metrics = yardstick::metric_set(rmse, mae, rsq),
  control = tune::control_grid(save_pred = TRUE)
)

@@

# Identify best penalty
tune::select_best(cv_results, metric = "rmse")
tune::select_best(cv_results, metric = "mae")
tune::select_best(cv_results, metric = "rsq")

best_penalty <- tune::select_best(cv_results, metric = "mae")

# Finalize Workflow with Best Penalty
final_wf <- tune::finalize_workflow(
  ridge_wf,
  best_penalty)

@@

# Set up Cross-Validation
folds <- rsample::group_vfold_cv(
  data_train_qb,
  group = gsis_id, # ensures all rows for a player are in the training set or all in the validation set for each fold
  v = 10) # 10-fold cross-validation

# Define Recipe (Formula)
rec <- recipes::recipe(
  fantasyPoints_lag ~ ., # use all predictors
  data = data_train_qb %>% select(-gsis_id, -fantasyPointsMC_lag))

# Define Model
enet_spec <- 
  parsnip::linear_reg(
    penalty = tune::tune(),
    mixture = tune::tune()) %>%
  parsnip::set_engine("glmnet")

@@

# Tune the Grid
cv_results <- tune::tune_grid(
  enet_wf,
  resamples = folds,
  grid = grid_enet,
  metrics = yardstick::metric_set(rmse, mae, rsq),
  control = tune::control_grid(save_pred = TRUE)
)

@@

# Identify best penalty
tune::select_best(cv_results, metric = "rmse")
tune::select_best(cv_results, metric = "mae")
tune::select_best(cv_results, metric = "rsq")

best_penalty <- tune::select_best(cv_results, metric = "mae")

# Finalize Workflow with Best Penalty
final_wf <- tune::finalize_workflow(
  enet_wf,
  best_penalty)

@@

# Set up Cross-Validation
folds <- rsample::group_vfold_cv(
  data_train_qb,
  group = gsis_id, # ensures all rows for a player are in the training set or all in the validation set for each fold
  v = 10) # 10-fold cross-validation

# Define Recipe (Formula)
rec <- recipes::recipe(
  fantasyPoints_lag ~ ., # use all predictors
  data = data_train_qb %>% select(-gsis_id, -fantasyPointsMC_lag))

# Define Model
rf_spec <- 
  parsnip::rand_forest(
    mtry = tune::tune(),
    min_n = tune::tune(),
    trees = 500) %>%
  parsnip::set_mode("regression") %>%
  parsnip::set_engine(
    "ranger",
    importance = "impurity")

@@

# Tune the Grid
cv_results <- tune::tune_grid(
  rf_wf,
  resamples = folds,
  grid = rf_grid,
  metrics = yardstick::metric_set(rmse, mae, rsq),
  control = tune::control_grid(save_pred = TRUE)
)

@@

# Identify best penalty
tune::select_best(cv_results, metric = "rmse")
tune::select_best(cv_results, metric = "mae")
tune::select_best(cv_results, metric = "rsq")

best_penalty <- tune::select_best(cv_results, metric = "mae")

# Finalize Workflow with Best Penalty
final_wf <- tune::finalize_workflow(
  rf_wf,
  best_penalty)

@@

smerf$forest # the fitted random forest (obtained at the last iteration)
smerf$random_effects %>% data.frame() # the predicted random effects for each player
smerf$omega %>% data.frame() # the predicted stochastic processes
smerf$OOB # OOB error at each iteration

# Predict on Test Data
predict(
  smerf,
  X = data_test_qb %>% as_tibble() %>% dplyr::select(season:att_br.rush) %>% as.matrix(),
  Z = data_test_qb[,c("pacr")] %>% as.matrix(),
  id = data_test_qb[,c("gsis_id")] %>% as.matrix(),
  time = data_test_qb[,c("ageCentered20")] %>% as.matrix())
```
